---
title: "strings and factors"
author: "Lorraine Kwok"
date: "October 15, 2019" 
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE)

library(tidyverse)
library(rvest)
library(p8105.datasets)
```

## String manipulation 

```{r}
string_vec = c("my", "name", "is", "jeff")

# find and tell me if string "jeff" exists
str_detect(string_vec, "jeff")

str_detect(string_vec, "m")
```

For str_detect, you can search for parts of the string or the entire string

str_replace(variable name, pattern, what you want the pattern to be replaced by)

```{r}
str_replace(string_vec, "jeff", "Jeff")

str_replace(string_vec, "m", "M")
```

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")

str_detect(string_vec, "^i think")
```

The ^ denotes starts with 
The $ denotes ends with 

We are interested in detecting words that start with a lowercase or capital b but not in all caps
```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec, "[Bb]ush")
```

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9][a-zA-Z]")
```

We're interested in identifying anything with a 7 followed by a character, followed by a 11. A "." matches any character at all.

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```

Sometimes regular expressions include brackets. What happens if you actually want to match a square bracket or the period instead of having it be treated as a placeholder. A backslash is also a special character - need two backslashes to search for [

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, pattern = "\\[")
```

## use strings in practice

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  select(id, visit, everything()) %>%
  mutate(
    visit = str_replace(visit, "bl", "00m"),
    visit = fct_relevel(visit, str_c(c("00", "01", "06", "12"), "m"))) %>%
  arrange(id, visit)
```

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) %>% 
  html_nodes(css = "table") %>% 
  .[[1]] %>%
  html_table() %>%
  slice(-1) %>%
  as_tibble()
```

```{r}
table_marj %>%
  select(-contains("P Value")) %>%
  pivot_longer(
    -State,
    names_to = "age_year",
    values_to = "percent"
  ) %>%
  separate(age_year, into = c("age", "year"), sep = "\\(") %>% 
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)
  )
```

## factors

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

fct_relevel: takes numeric value in the background and releveling it 
relevel can only specify the first level 

```{r}
weather_df %>%
  mutate(name = fct_relevel(name, c("Waikiki_HA", "CentralPark_NY", "Waterhole_WA"))) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

Reorder by tmax 

```{r}
weather_df %>%
  mutate(name = fct_reorder(name, tmax)) %>% 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

## airbnb data

```{r}
data("nyc_airbnb")

nyc_airbnb %>%
  filter(neighbourhood_group == "Manhattan") %>% 
  mutate(
    neighbourhood = fct_reorder(neighbourhood, price, na.rm = TRUE)) %>% 
  ggplot(aes(x = neighbourhood, y = price)) +
  geom_boxplot() +
  coord_flip() + 
  ylim(0, 1000)
```

