---
title: "strings and factors"
author: "Lorraine Kwok"
date: "October 15, 2019" 
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE)

library(tidyverse)
library(rvest)
library(p8105.datasets)
```

## String manipulation 

```{r}
string_vec = c("my", "name", "is", "jeff")

# find and tell me if string "jeff" exists
str_detect(string_vec, "jeff")

str_detect(string_vec, "m")
```

For str_detect, you can search for parts of the string or the entire string

str_replace(variable name, pattern, what you want the pattern to be replaced by)

```{r}
str_replace(string_vec, "jeff", "Jeff")

str_replace(string_vec, "m", "M")
```

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")

str_detect(string_vec, "^i think")
```

The ^ denotes starts with 
The $ denotes ends with 

We are interested in detecting words that start with a lowercase or capital b but not in all caps
```{r}
string_vec = c(
  "Y'all remember Pres. HW Bush?",
  "I saw a green bush",
  "BBQ and Bushwalking at Molonglo Gorge",
  "BUSH -- LIVE IN CONCERT!!"
  )

str_detect(string_vec, "[Bb]ush")
```

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9][a-zA-Z]")
```

We're interested in identifying anything with a 7 followed by a character, followed by a 11. A "." matches any character at all.

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11")
```

Sometimes regular expressions include brackets. What happens if you actually want to match a square bracket or the period instead of having it be treated as a placeholder. A backslash is also a special character - need two backslashes to search for []

```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, pattern = "\\[")
```

